---
title: "DTV - ITS2 analysis"
author: "Nicola G. Kriefall"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# SymPortal pre-submission

Submitting ITS2 samples to [Symportal](symportal.org)

Prior to submission: retaining only paired reads that match ITS2 primer (we pooled with 16S during sequencing)

```{bash setup symportal, eval=FALSE}
# *note: most of this was written by Dr. Carly D. Kenkel

# in Terminal home directory:
# following instructions of installing BBtools from https://jgi.doe.gov/data-and-tools/bbtools/bb-tools-user-guide/installation-guide/
# 1. download BBMap package, sftp to installation directory
# 2. untar: 
tar -xvzf BBMap_(version).tar.gz
# 3. test package:
cd bbmap
~/bin/bbmap/stats.sh in=~/bin/bbmap/resources/phix174_ill.ref.fa.gz

# primers for ITS2:
#fwd:
#GAATTGCAGAACTCCGTGAACC
#rev:
#CGGGTTCWCTTGTYTGACTTCATGC

# making a sample list based on the first phrase before the underscore in the .fastq name
ls *R1_001.fastq | cut -d '_' -f 2 > samples.list

# cuts off the extra words in the .fastq files
for file in $(cat samples.list); do mv *${file}*R1_001.fastq ${file}_R1.fastq; mv *${file}*R2_001.fastq ${file}_R2.fastq; done

# gets rid of reads that still have the adaptor sequence
# [I stopped doing this 'cause it never happened]
#for file in $(cat samples.list); do ~/bin/bbmap/bbduk.sh in1=${file}_R1.fastq in2=${file}_R2.fastq ref=adaptors.fasta out1=${file}_R1_NoIll.fastq out2=${file}_R2_NoIll.fastq; done &>bbduk_NoIll.log

##getting rid of first 4 bases (degenerate primers created them)
for file in $(cat samples.list); do ~/bin/bbmap/bbduk.sh in1=${file}_R1.fastq in2=${file}_R2.fastq ftl=4 out1=${file}_R1_No4N.fastq out2=${file}_R2_No4N.fastq; done &>bbduk_No4N.log

# only keeping reads that start with the primer
for file in $(cat samples.list); do ~/bin/bbmap/bbduk.sh in1=${file}_R1_No4N.fastq in2=${file}_R2_No4N.fastq k=15 restrictleft=25 literal=GAATTGCAGAACTCCGTGAACC,CGGGTTCWCTTGTYTGACTTCATGC outm1=${file}_R1_ITS.fastq outu1=${file}_R1_check.fastq outm2=${file}_R2_ITS.fastq outu2=${file}_R2_check.fastq; done &>bbduk_ITS.log
# higher k = more reads removed, but can't surpass k=20 or 21

# renamed them to the shorter version again after checking
for file in $(cat samples.list); do  mv ${file}_*R1*.fastq ${file}_R1.fastq; mv ${file}_*R2*.fastq ${file}_R2.fastq; done 
```

Then transferred all "R1.fastq" & "R2.fastq" files to the folder to be submitted to SymPortal

# ITS2 analysis setup

Information on Symportal output documents [here](https://github.com/didillysquat/SymPortal_framework/wiki/SymPortal-output-documents)

Formatting notes:

- Cleaned up file called 223_20220913T175714_DBV_20220914T011822.profiles.absolute.abund_and_meta.txt to format for phyloseq

## Setup

```{r packages setup, echo=FALSE}
library(phyloseq)
library('ggplot2')
library('Rmisc')
library(cowplot)
library("ggpubr")
library("vegan")
library(scales)
library(colorspace)

setwd("~/Documents/thermvar_microbes_master/thermvar_orbicella/tvm_its2")
```

## Checking for no read samples

```{r counts per sample, eval=F}
counts <- read.csv('dtv_symportal_type_profiles.csv',header=TRUE,row.names=1,check.names=FALSE)
#41 samples, 15 type profiles

sort(rowSums(counts))
counts.no0 <- counts[rowSums(counts)>0,]
#none failed
```

## Phyloseq object - type profiles

Ran once then saved to re-read in later

```{r phyloseq object type profiles, eval=F}
# import dataframe holding sample information
samdf <- read.csv("dtv_its2_sampleinfo_fixed.csv")
head(samdf)
rownames(samdf) <- samdf$sample_name

# import taxa info
taxa <- read.csv("dtv_symportal_taxa.csv",header=TRUE)
rownames(taxa) <- as.factor(taxa$ITS2_type_profile)
mtaxa <- as.matrix(taxa)

# import counts (absolute abundance from its2 type profiles)
mcounts <- as.matrix(counts)

# Construct phyloseq object 
ps <- phyloseq(otu_table(mcounts, taxa_are_rows=FALSE),
               sample_data(samdf),
               tax_table(mtaxa))

ps
# phyloseq-class experiment-level object
# otu_table()   OTU Table:         [ 15 taxa and 41 samples ]
# sample_data() Sample Data:       [ 41 samples by 8 sample variables ]
# tax_table()   Taxonomy Table:    [ 15 taxa by 4 taxonomic ranks ]

#saveRDS(ps,file="ps.its2.rds")
```

# Plotting

## Libs

```{r}
library("microshades")
library("DHARMa")
library("ggh4x")
```

## Re-read in data

```{r}
ps1 <- readRDS("ps.its2.RDS")
##remove redone sample
ps <- subset_samples(ps1, sample_name!="KA4-2")

ps.rel1 <- transform_sample_counts(ps, function(x) x / sum(x))

##removing duplicate clones
ps.rel <- subset_samples(ps.rel1,genotype_original!="KA" & genotype_original!="VB")
ps.rel
plot_bar(ps.rel,fill="ITS2_type_profile",x="genotype")

ps.fra <- subset_samples(ps.rel,host_species=="franksi")
ps.fav <- subset_samples(ps.rel,host_species=="faveolata")
```

## Microshades plot {.tabset}

### So much re-formatting

```{r}
mdf.ps1 <- prep_mdf(ps.rel,subgroup_level="Majority_ITS2_sequence")
mdf.ps.e <- subset(mdf.ps1,OTU=="E1c") ##checking out background levels
mdf.ps <- subset(mdf.ps1,OTU!="E1c") #removed

##reorder
mdf.ps$timepoint <- factor(mdf.ps$timepoint,levels=c("preheat","postheat"))

type_groups <- c(unique(mdf.ps$Majority_ITS2_sequence))

# Create a color object for the specified data
col.mdf.ps <- create_color_dfs(mdf.ps, top_orientation = FALSE,group_level="Clade",subgroup_level="Majority_ITS2_sequence",selected_groups=c("A","B","D"),cvd=TRUE)

#Extract
col.mdf.ps.m <- col.mdf.ps$mdf
col.mdf.ps.c <- col.mdf.ps$cdf

##renames
col.mdf.ps.m$treatment <- gsub("control", "Control",col.mdf.ps.m$treatment)
col.mdf.ps.m$treatment <- gsub("thermvar", "DTV",col.mdf.ps.m$treatment)
col.mdf.ps.m$timepoint <- gsub("preheat", "Pre-heat",col.mdf.ps.m$timepoint)
col.mdf.ps.m$timepoint <- gsub("postheat", "Post-heat",col.mdf.ps.m$timepoint)
col.mdf.ps.m$timepoint <- gsub("postheat", "Post-heat",col.mdf.ps.m$timepoint)
col.mdf.ps.m$timepoint <- gsub("postheat", "Post-heat",col.mdf.ps.m$timepoint)
col.mdf.ps.m$timepoint <- factor(col.mdf.ps.m$timepoint, levels=c("Pre-heat","Post-heat"))
col.mdf.ps.m$genotype <- gsub("KA_KG", "KA-KG",col.mdf.ps.m$genotype)
col.mdf.ps.m$genotype <- gsub("VB_VD", "VB-VD",col.mdf.ps.m$genotype)

col.mdf.ps.m.fav <- subset(col.mdf.ps.m,host_species=="faveolata")
col.mdf.ps.m.fra <- subset(col.mdf.ps.m,host_species=="franksi")
```

### The plots

```{r}
##fav
gg.its2.fav <- ggplot(col.mdf.ps.m.fav,aes(x=genotype,y=Abundance,fill=Majority_ITS2_sequence))+
  geom_bar(stat="identity")+
  facet_grid(timepoint~treatment)+
  theme_bw()+
  xlab("Coral genet")+
  ylab("Relative abundance")+
  ggtitle(expression(italic("O. faveolata")))+
  scale_fill_manual(values=c("#616161","#4E7705","#97CE2F","#098BD9"),name="Majority ITS2 seq.")+
  theme(legend.position="none")
gg.its2.fav

##franks
gg.its2.fra <- ggplot(col.mdf.ps.m.fra,aes(x=genotype,y=Abundance,fill=Majority_ITS2_sequence))+
  geom_bar(stat="identity")+
  facet_grid(timepoint~treatment)+
  theme_bw()+
  xlab("Coral genet")+
  ylab(" ")+
  ggtitle(expression(italic("O. franksi")))+
  scale_fill_manual(values=c("#616161","#4E7705","#97CE2F","#DDFFA0","#098BD9"),name="Majority ITS2 seq.")

gg.its2.both <- ggarrange(gg.its2.fav,gg.its2.fra,labels="AUTO",legend="none")

##I want the franks legend
legend <- get_legend(gg.its2.fra)

gg.its2.multi <- plot_grid(
  gg.its2.both, legend, ncol = 2,
  rel_widths = c(1, 0.2)
)
gg.its2.multi
#ggsave("gg.its2.revised.png",width=8,height=3)
```

# Mixed models

## Setup

```{r}
##retaining sample sums
otu.all <- data.frame(ps.rel@otu_table)
sam.all <- data.frame(ps.rel@sam_data)
row.names(otu.all)==row.names(sam.all)
sam.all$counts.sum <- rowSums(otu.all)
ps.rel@sam_data <- sample_data(sam.all)

ps.glom <- tax_glom(ps.rel,"Majority_ITS2_sequence")
ps.glom
```

## Trim, thoroughly

```{r}
table(ps.glom@sam_data$treatment)
table(ps.glom@sam_data$timepoint)
ps.glom@tax_table

##percentage that includes at least 3 samples
3/32

ps.glom.trim1 <- filter_taxa(ps.glom, function(x) sum(x > 0) > (0.09*length(x)), TRUE)
ps.glom.trim1 ##5 taxa, 32 samples

##any samples missing data
ps.glom.trim <- prune_samples(sample_sums(ps.glom.trim1) > 0, ps.glom.trim1)
ps.glom.trim #none missing

ps.glom.trim@tax_table

##proportion of zeroes after
otu.trim <- data.frame(ps.glom.trim@otu_table)
sum(otu.trim == 0) / (ncol(otu.trim) * nrow(otu.trim)) * 100
#68.125

##before
otu.glom <- data.frame(ps.glom@otu_table)
sum(otu.glom == 0) / (ncol(otu.glom) * nrow(otu.glom)) * 100
#72.91667
```

### More formatting - long form

```{r}
#create sample name column for joining
otu.trim$sample_name <- row.names(otu.trim)

#join the data sets by sample name
otu.tot1 <- plyr::join(otu.trim,sam.all, by="sample_name", type="left")
colnames(otu.tot1)

##getting rid of some irrelevant metadata
otu.tot <- otu.tot1 %>%
  dplyr::select(-genotype_original,-fragment,-treat_time,-counts.sum)
colnames(otu.tot)

tot.long <- reshape::melt(otu.tot, id.vars=c("sample_name", "host_species","treatment","timepoint", "genotype"))

##not sure if I need to do these:
#add unique id for each row in the data. useful to control for overdispersion in models that assume error distributions that cannot account for it
# sid.long$rowid <- 1:nrow(sid.long)
# 
# sid.long$value.pa <- sid.long$value
# sid.long$value.pa[sid.long$value.pa>0] <- 1

#saveRDS(sid.long,file="sid.long.mixmodtrim.rds")
```

# Running the statz

- variable = ITS2 type
- value = the counts for the ITS2 type

```{r}
library("glmmTMB")

##option:
#control=glmmTMBControl(profile=TRUE)
mod.tot.ord <- glmmTMB(value~timepoint*treatment+host_species+(1|genotype)+(1|variable)+(1|variable:genotype)+(1|variable:treatment)+(1|variable:timepoint)+(1|variable:host_species)+(1|variable:timepoint:treatment), family=ordbeta, data=tot.long)
summary(mod.tot.ord)

mod.tot.ord.resid <- simulateResiduals(fittedModel = mod.tot.ord, plot = T)
##actually not bad

mod.tot.ord.noint <- update(mod.tot.ord,.~.-(1|variable:timepoint:treatment))
anova(mod.tot.ord,mod.tot.ord.noint) #ns

mod.tot.ord.trt <- update(mod.tot.ord.noint,.~.-(1|variable:treatment))
anova(mod.tot.ord.noint,mod.tot.ord.trt) #ns

mod.tot.ord.tim <- update(mod.tot.ord.noint,.~.-(1|variable:timepoint))
anova(mod.tot.ord.noint,mod.tot.ord.tim) #sig *0.05

mod.tot.ord.hos <- update(mod.tot.ord.noint,.~.-(1|variable:host_species))
anova(mod.tot.ord.noint,mod.tot.ord.hos) #almost sig 0.09 .
```

## ggpredict

```{r}
library("ggeffects")

pred.tot.ord <- ggpredict(mod.tot.ord.noint,terms=c("variable","timepoint"),type="random")

plot(pred.tot.ord)

pred.tot.ord$group <- factor(pred.tot.ord$group,levels=c("preheat","postheat"))

ggplot(pred.tot.ord,aes(x=group,y=predicted,fill=x))+
  geom_bar(stat="identity",position=position_dodge(1))+
  #geom_errorbar(aes(ymin=predicted-conf.high,ymax=predicted+conf.high),width=0.1)+
  facet_wrap(~x,scales="free")
```

## Back to the actual data

```{r}
##the actual data
library(dplyr)

ps.glom.trim.melt <- psmelt(ps.glom.trim)

avg_abund_by_treatment <- ps.glom.trim.melt %>%
  group_by(timepoint, Majority_ITS2_sequence) %>%
  summarise(mean_abundance = mean(Abundance))

avg_abund_by_treatment <- ps.glom.trim.melt %>%
  group_by(timepoint, Majority_ITS2_sequence) %>%
  summarise(
    mean_abundance = mean(Abundance, na.rm = TRUE),
    se_abundance = sd(Abundance, na.rm = TRUE) / sqrt(n())
  )

avg_abund_by_treatment$timepoint <- factor(avg_abund_by_treatment$timepoint,levels=c("preheat","postheat"))

gg.avg <- ggplot(avg_abund_by_treatment,aes(x=timepoint,y=mean_abundance,fill=Majority_ITS2_sequence))+
  geom_bar(stat="identity",color="black")+
  facet_wrap(~Majority_ITS2_sequence,ncol=5)+
  geom_errorbar(aes(ymin=mean_abundance-se_abundance,ymax=mean_abundance+se_abundance),width=0.2)+
  scale_fill_manual(values=c("#616161","#4E7705","#97CE2F","#DDFFA0","#098BD9"),name="Majority ITS2 seq.")+
  theme_bw()+
  ylab("Mean rel. abundance")+
  xlab("")+
  scale_x_discrete(labels=c("Pre-heat","Post-heat"))+
  theme(axis.text.x = element_text(angle=45,hjust=1),legend.position="none",plot.margin=margin(t=5,r=100,b=5,l=5))
gg.avg
```

# Figure 3

```{r}
ggarrange(gg.its2.multi,gg.avg,nrow=2,heights=c(1,0.8),labels=c("","C"))

#ggsave("gg.its2.revised.multi.png",width=8,height=6)
#ggsave("gg.its2.revised.multi.pdf",width=8,height=6)
```
