---
title: "Temperature data"
author: "Nicola G. Kriefall"
date: "6/30/2021"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
---

# Setup
setwd("~/Desktop/BU NSF Postdoc/Nicola_Ofav_chpt3/GitHub Versions of Code")

```{r not public, include=FALSE}
#setwd("/Volumes/GoogleDrive/My Drive/thermvar_microbes_master/thermvar_orbicella/temp_logs")
setwd("/Volumes/GoogleDrive-104519233854090018057/My Drive/thermvar_microbes_master/thermvar_orbicella/temp_logs")
```

```{r libraries}
library(ggplot2)
library(lubridate)
library(cowplot)
library(dplyr)
library(car)
library(bestNormalize)
library(coin)
```

```{r data}
temp.all <- read.csv("hobo_all.csv")
```

## Date cleaning

```{r}
str(temp.all)

temp.all$datetime <- mdy_hm(temp.all$Date.Time..EST.,tz=Sys.timezone())

#me trying to get R to understand my date formatting
temp.all$date <- paste(temp.all$year,"/",temp.all$mon,"/",temp.all$day)
temp.all$date <- gsub(' ', '', temp.all$date)
temp.all$date <- as.Date(temp.all$date,format="%Y/%m/%d")
str(temp.all)

temp.all$treatment <- factor(temp.all$treatment,levels=c("var","con"))
```

# Processing & plotting data

## Plot with all data

```{r}
ggplot(data=temp.all,aes(x=log_number,y=temp,color=treatment,group=treatment))+
  geom_line()+
  scale_color_manual(values=c("#D66982","#B9C184"),labels=c("TV","Ctrl"))+
  theme_bw()

#ggsave("temp.all.pdf")
```

## Plot with all data minus the control after heat ramping

Remove the dip in temp when control logger removed from tanks due to tank consolidations
```{r}
# Confirm the exact timestamp when the ctrl logger removed because all corals were moved to the same tank for heat expt.
temp.all %>%
  filter(treatment == "con", temp < 23) %>%
  summarise(dip_start = min(log_number))

#consolidation_time <- ymd_hm("2020-XX-XX HH:MM", tz = Sys.timezone())  # <- replace with value from above
# 
# temp.clean <- temp.all %>%
#   filter(!(treatment == "con" & datetime >= consolidation_time))

dip_start <- temp.all %>%
  filter(treatment == "con", temp < 23) %>%
  summarise(dip_start = min(log_number)) %>%
  pull(dip_start)

temp.clean <- temp.all %>%
  filter(!(treatment == "con" & log_number >= 10850))
```
Plot clean figure without dip
```{r}
ggplot(data=temp.clean,aes(x=datetime,y=temp,color=treatment,group=treatment))+
  geom_line()+
  scale_color_manual(values=c("#E24A13","#2C2673"),labels=c("TV","Ctrl"))+
  theme_bw()

#ggsave("temp.all.pdf", width = 7, height = 5, units = "in")
```


## Data subset - just 90 day acclimation

```{r}
temp.90 <- subset(temp.all,log_number>=760&log_number<=9844) #should be approximately 10am on 8/21, then midnight on 11/24 (might update later)
ggplot(data=temp.90,aes(x=datetime,y=temp,color=treatment,group=treatment,linetype=treatment))+
  scale_linetype_manual(name="Treat",values=c("twodash","solid"),labels=c("TV","Ctrl"))+
  geom_line()+
  #geom_hline(yintercept=25.67,linetype=2)+
  theme_bw()+
  scale_color_manual(name="Treat",values=c("#D66982","#B9C184"),labels=c("TV","Ctrl"))+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  ylab("Temperature (C)")+
  xlab("")
  #scale_linetype_manual(values=c("solid","dotted"))

#ggsave("tvm.temp.90d.pdf",height=2,width=6)

##30 days instead..
temp.30 <- subset(temp.all,log_number>=1775&log_number<4659) #should be approximately 10am on 8/21, then midnight on 11/24 (might update later)
ggplot(data=temp.30,aes(x=datetime,y=temp,color=treatment,group=treatment,linetype=treatment))+
  scale_linetype_manual(name="Treat",values=c("twodash","solid"),labels=c("TV","Ctrl"))+
  geom_line()+
  #geom_hline(yintercept=25.67,linetype=2)+
  theme_bw()+
  scale_color_manual(name="Treat",values=c("#D66982","#B9C184"),labels=c("TV","Ctrl"))+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  ylab("Temperature (C)")+
  xlab("Sep - Nov")
  #scale_linetype_manual(values=c("solid","dotted"))

#ggsave("tvm.temp.30d.pdf",height=2,width=4)
```

### Summary stats

```{r}
leveneTest(temp~treatment,data=temp.90)
wilcox.test(temp~treatment,data=temp.90)
temp.90 %>% wilcox_effsize(temp ~ treatment)

temp.90.con <- subset(temp.90,treatment=="con")
mean(temp.90.con$temp,na.rm=T)
sd(temp.90.con$temp,na.rm=T)

temp.90.var <- subset(temp.90,treatment=="var")
mean(temp.90.var$temp,na.rm=T)
sd(temp.90.var$temp,na.rm=T)
```

### Daily means

```{r}
temp.90.means <- temp.90 %>%
  group_by(date,treatment) %>%
  summarize_at('temp',mean,na.rm=T)

ggplot(temp.90.means,aes(x=date,y=temp,color=treatment))+
  geom_boxplot()+
  theme_bw()

leveneTest(temp~treatment,data=temp.90.means) #ns
shapiro.test(log(temp.90.means$temp)) #very sig

temp.norm <- orderNorm(temp.90.means$temp)
temp.90.means$temp.norm <- temp.norm$x.t
shapiro.test(temp.90.means$temp.norm) #ns

a1 <- aov(temp.norm~treatment,data=temp.90.means)
summary(a1)

wilcox.test(temp~treatment,data=temp.90.means)
```

## Thermal Variability
```{r}
daily_temp_range <- temp.90 %>%
  filter(treatment %in% c("con", "var")) %>%
  group_by(treatment, date) %>%
  summarise(
    daily_range = max(temp, na.rm = TRUE) -
                  min(temp, na.rm = TRUE),
    n = sum(!is.na(temp)),
    .groups = "drop"
  ) %>%
  filter(n > 1)   # ensures a real range

temp_range_summary <- daily_temp_range %>%
  group_by(treatment) %>%
  summarise(
    mean_daily_range = mean(daily_range, na.rm = TRUE),
    se_daily_range   = sd(daily_range, na.rm = TRUE) /
                       sqrt(sum(!is.na(daily_range))),
    .groups = "drop"
  )
temp_range_summary
```

## Data subset - just ramping up

```{r}
temp.ramp <- subset(temp.all,log_number<760) 
ggplot(data=temp.ramp,aes(x=datetime,y=temp,color=treatment,group=treatment,linetype=treatment))+
  geom_line()+
  #geom_hline(yintercept=25.67,linetype=2)+
  theme_bw()+
  scale_color_manual(name="Treat",values=c("#D66982","#B9C184"),labels=c("TV","Ctrl"))+
  scale_linetype_manual(name="Treat",values=c("twodash","solid"),labels=c("TV","Ctrl"))+
  ylab("Temperature (C)")+
  xlab("")+
  theme(axis.text.x=element_text(angle=45,hjust=1))

#ggsave("tvm.temp.ramp.pdf",height=2,width=3)
```

## Data subset - heat

```{r}
temp.heat <- subset(temp.all,log_number>9889) 
temp.heat.tv <- subset(temp.heat,treatment=="var") 
ggplot(data=temp.heat.tv,aes(x=datetime,y=temp))+
  geom_line(color="pink",linetype="solid")+
  geom_line(color="black",linetype="dotted")+
  #geom_hline(yintercept=25.67,linetype=2)+
  theme_bw()+
  #scale_color_manual(name="",values=c("#ECA313"),labels=c("Heat"))+
  ylab("Temperature (C)")+
  xlab("")+
  #scale_linetype_manual(name="",values="dotted",labels=c("Heat"))+
  theme(axis.text.x=element_text(angle=45,hjust=1))

#ggsave("tvm.temp.heat.pdf",height=2,width=3)
```

Averages per treatment

```{r}
temp.var <- subset(temp.90,treatment=="var")
summary(temp.var$temp) #25.73

temp.con <- subset(temp.90,treatment=="con")
summary(temp.con$temp) #25.67
```

# Other water quality

## Setup things

```{r}
library(lubridate)
library(ggplot2)
library("Rmisc")
library("car")
```

```{r}
#setwd("/Volumes/GoogleDrive-104519233854090018057/My Drive/thermvar_microbes_master/thermvar_orbicella/temp_logs")
setwd("/Volumes/GoogleDrive/My Drive/thermvar_microbes_master/thermvar_orbicella/temp_logs")

wq <- read.csv("Tank_WQ - Sheet1.csv")
str(wq)
wq$date <- mdy(wq$date)
str(wq)
```

## Plot temp

Not really helpful since no measurements at night

```{r}
ggplot(wq,aes(x=date,y=temperature,color=tank))+
  geom_point()+
  theme_bw()
```

## Salinity

### Salinity plot

```{r}
# ggplot(wq,aes(x=date,y=salinity,color=tank))+
#   geom_point()+
#   theme_bw()+
#   geom_smooth()

ggplot(wq,aes(x=tank,y=salinity,color=tank))+
  geom_boxplot()+
  theme_bw()
```

### Salinity stats

```{r}
leveneTest(salinity~tank,data=wq) #ns
shapiro.test(wq$salinity) #ns

a.sal <- aov(salinity~tank,data=wq)
summary(a.sal) #very not significant!
```

## Other nutrients

```{r}
wq <- subset(wq,tank!="Sys heat")
```

### Nitrate

```{r}
ggplot(wq,aes(x=tank,y=nitrate,color=tank))+
  geom_boxplot()

##all 0, no stats
```

### Phosphate

```{r}
ggplot(wq,aes(x=tank,y=phosphate,color=tank))+
  geom_boxplot()

leveneTest(phosphate~tank,data=wq) #ns
shapiro.test(wq$phosphate) #sig

wilcox.test(phosphate~tank,data=wq) #ns
```

### Calcium

```{r}
ggplot(wq,aes(x=tank,y=calcium,color=tank))+
  geom_boxplot()

leveneTest(calcium~tank,data=wq) #ns
shapiro.test(wq$calcium) #sig

wilcox.test(calcium~tank,data=wq) #ns
```

### Alkalinity

```{r}
ggplot(wq,aes(x=tank,y=alkalinity,color=tank))+
  geom_boxplot()

leveneTest(alkalinity~tank,data=wq) #ns
shapiro.test(wq$alkalinity) #almost sig

wilcox.test(alkalinity~tank,data=wq) #ns
```

### Magnesium

```{r}
ggplot(wq,aes(x=tank,y=magnesium,color=tank))+
  geom_boxplot()

leveneTest(magnesium~tank,data=wq) #ns
shapiro.test(wq$magnesium) #ns

wilcox.test(magnesium~tank,data=wq) #ns
```

### Summary stats

```{r}
wq.co <- subset(wq,tank=="Sys 6")
wq.tv <- subset(wq,tank=="Sys 8")

##salinity
mean(wq.co$salinity,na.rm=TRUE)
sd(wq.co$salinity,na.rm=TRUE)

mean(wq.tv$salinity,na.rm=TRUE)
sd(wq.tv$salinity,na.rm=TRUE)

##nitrate
mean(wq.co$nitrate,na.rm=TRUE)
sd(wq.co$nitrate,na.rm=TRUE)

mean(wq.tv$nitrate,na.rm=TRUE)
sd(wq.tv$nitrate,na.rm=TRUE)

##phosphate
mean(wq.co$phosphate,na.rm=TRUE)
sd(wq.co$phosphate,na.rm=TRUE)

mean(wq.tv$phosphate,na.rm=TRUE)
sd(wq.tv$phosphate,na.rm=TRUE)

##calcium
mean(wq.co$calcium,na.rm=TRUE)
sd(wq.co$calcium,na.rm=TRUE)

mean(wq.tv$calcium,na.rm=TRUE)
sd(wq.tv$calcium,na.rm=TRUE)

##alkalinity
mean(wq.co$alkalinity,na.rm=TRUE)
sd(wq.co$alkalinity,na.rm=TRUE)

mean(wq.tv$alkalinity,na.rm=TRUE)
sd(wq.tv$alkalinity,na.rm=TRUE)

##magnesium
mean(wq.co$magnesium,na.rm=TRUE)
sd(wq.co$magnesium,na.rm=TRUE)

mean(wq.tv$magnesium,na.rm=TRUE)
sd(wq.tv$magnesium,na.rm=TRUE)


```


